#I - Loading and preprocessing data
##1. Load the data
```{r}
library(ggplot2)
library(plyr)
activity <- read.csv("C:\\Users\\tuuye\\AppData\\Local\\Temp\\Rar$DR08.934\\activity.csv")
```
##2. Process the data
```{r}
activity$day <- weekdays(as.Date(activity$date))
activity$DateTime <- as.POSIXct(activity$date, format="%Y-%m-%d")
clean <- activity[!is.na(activity$steps),]
```

#II - What is mean total number of steps taken per day?
##1. Calculate the total number of steps taken per day
```{r}
sumTable <- aggregate(activity$steps ~ activity$date, FUN=sum,)
colnames(sumTable) <- c("Date", "Steps")
```
##2. Make a histogram of the total number of steps taken each day
```{r}
hist(sumTable$Steps, breaks=5, xlab="Steps", main = "Total Steps per day", col = "purple")
```

## 3. Calculate and report the mean and median of the total number of steps taken perday
### - Mean of steps
```{r}
stepsmean <- as.integer(mean(sumTable$Steps))
print(stepsmean)
```
### - Median of steps
```{r}
stepsmedian <- as.integer(median(sumTable$Steps))
print(stepsmedian)
```

# III - What is the average daily activity pattern?
## 1. Make a time series plot
```{r}
library(plyr)
library(ggplot2)
clean <- activity[!is.na(activity$steps),]
interval_Steps <- ddply(clean, .(interval), summarize, Avg = mean(steps))
p <- ggplot(interval_Steps, aes(x=interval, y=Avg), xlab = "Time Interval", ylab="Average Number of Steps")
p + geom_line()+xlab("Interval")+ylab("Average Number of Steps")+ggtitle("Time Series Plot of the Average Number of Steps per Interval")
```

## 2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
maxSteps <- max(interval_Steps$Avg)
interval_Steps[interval_Steps$Avg==maxSteps,1]
```

# IV - Imputing missing values
## 1. Calculate and report the total number of missing values in the dataset
```{r}
nrow(activity[is.na(activity$steps),])
```

## 2. Devise a strategy for filling in all of the missing values in the dataset
```{r}
## Average number of steps per weekday and interval
data1 <- ddply(clean, .(interval, day), summarize, Avg = mean(steps))
## Dataset with all NAs for substitution
data_NA<- activity[is.na(activity$steps),]
## Merge 
merged_data<-merge(data_NA, data1, by=c("interval", "day"))
```

## 3. Create a new dataset that is equal to the original dataset but with the missing data filled in
```{r}
new_dataset<- merged_data[,c(6,4,1,2,5)]
colnames(new_dataset)<- c("steps", "date", "interval", "day", "DateTime")
data_combine <- rbind(clean, new_dataset)
```

## 4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day
```{r}
##Sum of steps
sum_steps <- aggregate(data_combine$steps ~ data_combine$date, FUN=sum)
colnames(sum_steps)<- c("Date", "Steps")

## Mean of the total number of steps taken per day
meansteps <- as.integer(mean(sum_steps$Steps))
print(meansteps)
## Median of the total number of steps taken per day
mediansteps <- as.integer(median(sum_steps$Steps))
print(mediansteps)

## Make a histogram
hist(sum_steps$Steps, breaks=5, xlab="Steps", main = "Total Steps per Day with NAs Fixed", col="Blue")

##The overall shape of the distribution has not change even though the new mean and median increased 
```
# V - Are there differences in activity patterns between weekdays and weekends?
```{r}
data_combine$DayCategory <- ifelse(data_combine$day %in% c("Saturday", "Sunday"), "Weekend", "Weekday")
library(lattice) 
intervalTable <- ddply(data_combine, .(interval, DayCategory), summarize, Avg = mean(steps))
xyplot(Avg~interval|DayCategory, data=intervalTable, type="l",  layout = c(1,2),
       main="Average Steps per Interval Based on Type of Day", 
       ylab="Average Number of Steps", xlab="Interval")
```

     